Type Classes for Mathematics in Type Theory

### 2010-12-10 papers/mscs/mscs.tex

Because the Coq type theory lacks quotient types (as it would make type checking undecidable), one usually bases abstract structures on a \emph{setoid} (`Bishop set'): a type equipped with an equivalence relation~\cite{Bishop67,Hofmann,Capretta}. As we will see in section~\ref{univ}, this way of working pays off when working with notions such as quotient algebras.

@@

Type classes have been a great success story in the Haskell functional programming language, as a means of organizing interfaces of abstract structures. Coq's type classes provide a superset of their functionality, but implemented in a different way.

@@

>>>>

Because the Coq type theory lacks quotient types (as it would make type checking undecidable), one usually bases abstract structures on a \emph{setoid} (`Bishop set'): a type equipped with an equivalence relation~\cite{Bishop67,Hofmann,Capretta}. Palmgren~\cite{palmgren2009constructivist} shows that Bishop sets have pleasant categorical properties, which translate to a powerful implicit type structure. It would be of interest to actually provide machine support for this type structure. As we will see in section~\ref{univ}, working with setoids pays off when working with notions such as quotient algebras.

@@

Type classes \cite{wadler1989adhoc} have been a great success story in the Haskell functional programming language, as a means of organizing interfaces of abstract structures. Coq's type classes provide a superset of their functionality, but implemented in a different way.

@@

The basic idea of using type-class-like facilities for structuring computerized mathematics dates back to the \textsc{AXIOM} computer algebra system~\cite{jenks1992axiom}. Weber~\cite{weber1993type} pursues the analogy between \textsc{AXIOM}'s so-called categories and type classes in Haskell. Santas~\cite{santas1995type} pursues analogies between type classes, \textsc{AXIOM} categories and existential types. Existential types are present in Haskell, but absent from Coq.
